\documentclass[10pt,twocolumn]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{bbm}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!70!black,
  citecolor=blue!70!black,
  urlcolor=blue!70!black,
}

% -----------------------------------------------------------------------
\title{\textbf{CTDenoiser}: When Fidelity Gains Erode Detectability\\
       A Benchmark for Low-Dose CT Denoising Across Supervised
       and Self-Supervised Regimes}

\author{%
  Tim Sereda\\
  \texttt{timgsereda@gmail.com}
}
\date{June 2026}
% -----------------------------------------------------------------------

\begin{document}
\maketitle

% -----------------------------------------------------------------------
\begin{abstract}
Low-dose computed tomography (LDCT) denoising must contend with structured,
spatially correlated noise that ordinary blind-spot self-supervision is not
designed to handle.  We present \textbf{CTDenoiser}, an open benchmark and
unified data, training, and inference harness that evaluates five
architectures---RED-CNN, DnCNN, U-Net, CTformer, and a Conditional Flow
Matching (CFM) model---across the supervision regimes for which they are
defined, from fully supervised training to similarity-based, blind-spot, and
per-image self-supervision.  On the Mayo TCIA abdomen data, supervised models
recover $\sim$4\,dB PSNR over the noisy input and finish within $0.1$\,dB of one
another across CNN backbones, so the supervision signal matters more than the
architecture.  Among label-free methods, only the correlated-noise-aware
Noise2Sim reliably beats the noisy baseline while blind-spot Noise2Void is
flat-to-negative---direct evidence that CT's correlated noise violates the
pixel-independence assumption these methods rely on.  Motivated by the resulting
$3$--$4$\,dB gap, we observe that the velocity field of an \emph{unconditional}
rectified flow, trained on two noisy observations of the same signal, recovers
the Noise2Noise posterior mean in a \emph{single} step with no clean targets.
We instantiate this as SSFlow, pairing the flow with a similarity-based
construction that decorrelates CT noise; it attains $+1.39$\,dB, the best
label-free result in our benchmark.  Finally, a task-based evaluation with a
channelized Hotelling observer shows these fidelity gains do not translate into
detectability: the same models that gain $\sim$4\,dB PSNR \emph{erode}
low-contrast lesion detectability rather than improving it---without fabricating
structure---a shift PSNR and SSIM do not register.  All code, model definitions,
and data-loading utilities are released at
\href{https://github.com/tsereda/CTDenoiser}{github.com/tsereda/CTDenoiser}.
\end{abstract}

% -----------------------------------------------------------------------
\section{Introduction}

Computed tomography (CT) is a cornerstone of modern medical imaging, but
the ionising radiation dose required for high-quality scans poses a
non-trivial lifetime cancer risk~\cite{brenner2007computed}.  Clinical
practice therefore frequently acquires low-dose CT (LDCT) scans and
applies post-hoc denoising.  The resulting reconstruction quality is
constrained by a fundamental trade-off: aggressive denoising suppresses
quantum noise but blurs fine anatomical structures; insufficient
denoising preserves sharpness but leaves noise that hampers radiological
interpretation.

Deep learning has transformed LDCT denoising.  Convolutional
encoder-decoders~\cite{chen2017lowdose} and
transformers~\cite{wang2022ctformer} now achieve near full-dose image
quality on the TCIA LDCT benchmark~\cite{moen2021low}, while diffusion
and flow-based generative models are beginning to emerge as an
alternative that models the full \emph{distribution} of plausible clean
images rather than a single point estimate.  The harder regime is
\emph{label-free} denoising, where paired full-dose targets are
unavailable; here CT's spatially correlated reconstruction noise breaks the
pixel-independence assumption that underpins blind-spot and downsampling
self-supervision.

Despite this progress, comparing methods fairly is difficult: published
results use heterogeneous preprocessing pipelines, evaluation patches,
normalisation conventions, and metric implementations~\cite{eulig2024benchmark}.
We address this with CTDenoiser, a self-contained benchmark that:

\begin{itemize}
  \item provides a single training loop and evaluation harness usable
        without modification for all five included architectures and five
        supervision regimes (supervised, similarity-based and blind-spot
        self-supervised, and two zero-shot per-image methods);
  \item enforces patient-level train/validation splits to prevent slice
        leakage;
  \item evaluates on full CT slices via overlapped tiling rather than
        held-out patches;
  \item augments the standard pixel-fidelity metrics with GMSD and
        a Residual Spectral Ratio (RSR) to surface the perceptual
        trade-offs of generative denoisers.
\end{itemize}

Our contributions are: (i) the harness and a corrected full-slice
evaluation (Section~\ref{sec:inference}); (ii) a controlled comparison
showing supervised denoisers gain $\sim$4\,dB over the noisy input while
the only label-free method to reliably beat it is the correlated-noise-aware
Noise2Sim (Section~\ref{sec:results}); (iii) the observation that an
\emph{unconditional} rectified flow recovers the one-step Noise2Noise
optimum without clean targets, together with its empirical instantiation
SSFlow that targets the remaining gap (Section~\ref{sec:method}); and
(iv) a task-based, hallucination-aware detectability evaluation---a channelized
Hotelling observer with a corrected noise-power spectrum---showing the
benchmark's PSNR/SSIM gains do not translate into low-contrast lesion
detectability (Section~\ref{sec:detect}).

% -----------------------------------------------------------------------
\section{Related Work}

\paragraph{CNN-based denoisers.}
RED-CNN~\cite{chen2017lowdose} introduced a residual encoder-decoder
with multi-level skip connections and remains a standard baseline.
DnCNN~\cite{zhang2017beyond} demonstrated that a flat stack of
Conv-BN-ReLU layers with a single residual prediction (noise subtraction
rather than clean-image prediction) achieves strong performance on
natural image denoising and transfers well to CT.

\paragraph{Transformer denoisers.}
CTformer~\cite{wang2022ctformer} applies a U-shaped transformer with
Token2Token Dilation (T2TD) blocks that aggregate local neighbourhood
tokens via dilated convolutions interleaved with cyclic spatial shifts.
This allows long-range self-attention without the quadratic memory cost
of global attention on full-resolution feature maps.

\paragraph{Generative approaches.}
Score-based diffusion models~\cite{song2021scorebased} have been applied
to CT denoising~\cite{gao2023corediff} and can produce perceptually
superior results, but inference requires hundreds to thousands of neural
function evaluations (NFEs).  Conditional flow matching~\cite{lipman2022flow,
albergo2022building} reduces this to as few as 10--50 NFEs by learning
straight-line probability paths between source and target distributions,
making it practical for clinical deployment.  Recent few-/single-step
generative denoisers push this further: Poisson Flow Consistency Models
(PFCM)~\cite{hein2025pfcm} distil a generative LDCT denoiser to a single step
and report that an unconstrained generative process introduces substantial
variability across denoised images, motivating fidelity-preserving
sampling---a concern our self-supervised flow addresses by construction
(Section~\ref{sec:method}).  More fundamentally, the perception--distortion
trade-off~\cite{blau2018perception} establishes that minimising distortion
necessarily costs perceptual realism for any distortion measure, the formal
reason a generative denoiser can look realistic while being unfaithful.

\paragraph{Self-supervised and zero-shot denoisers.}
When paired full-dose targets are unavailable, denoisers can be trained
from noisy data alone.  Noise2Noise~\cite{lehtinen2018noise2noise} showed
that mapping between two independent noisy observations of the same scene
recovers the clean signal in expectation.  Noise2Void~\cite{krull2019noise2void}
removes the need for a second observation via a blind-spot scheme that
withholds each target pixel from the network's receptive field, but it
assumes pixel-wise \emph{independent} noise---an assumption violated by CT,
whose reconstruction noise is spatially correlated.
Neighbor2Neighbor~\cite{huang2021neighbor2neighbor} manufactures a
Noise2Noise pair from a single image by sub-sampling neighbouring pixels,
inheriting the same independence assumption.  Two lines of work address
correlated noise directly: Noise2Sim~\cite{niu2020noise2sim} learns to map
between the centre pixels of self-similar non-local patches, suppressing
correlated as well as independent noise and remaining asymptotically
equivalent to supervised learning; and Zero-Shot Noise2Noise
(ZS-N2N)~\cite{mansour2023zsn2n} trains a tiny network per image from a
downsampled Noise2Noise pair, requiring no external data, no noise model,
and no pretraining.  In the vision literature, asymmetric pixel-shuffle
downsampling (AP-BSN)~\cite{lee2022apbsn} breaks spatial correlation with a
large training stride while preserving detail with a small inference stride,
the mechanism most correlated-noise blind-spot successors build on.  More
recently, Filter2Noise~\cite{sun2025filter2noise} couples self-supervised
single-image training with an interpretable attention-guided bilateral
filter designed specifically for LDCT, and WIA-LD2ND~\cite{zhao2024wia}
performs self-supervised LDCT denoising in the wavelet domain.  Our
benchmark includes Noise2Void, ZS-N2N, and Filter2Noise as label-free
baselines and adds Noise2Sim as the correlated-noise-aware extension.

\paragraph{Task-based and hallucination-aware evaluation.}
\label{sec:rw-taskbased}
Pixel-fidelity metrics such as PSNR and SSIM are increasingly recognised as
poor proxies for diagnostic quality on medical images, both because they
misrank novel restoration algorithms~\cite{breger2025reassess} and because
they do not track radiologist perception of denoised
CT~\cite{lee2025ldctiqa}.  The more rigorous standard is \emph{task-based}
assessment: model observers such as the Channelized Hotelling Observer (CHO)
estimate detectability ($d'$, AUC) on signal-known-exactly /
background-known-statistically tasks.  Bhadra et al.~\cite{bhadra2021halluc}
gave the first formal definition of hallucination in tomographic
reconstruction via the measurement / null-space decomposition, and Tivnan et
al.~\cite{tivnan2024halluc} introduced a Hallucination Index for generative
reconstruction that distinguishes added noise from added structure.  Applied
to denoisers specifically, Li et al.~\cite{li2021detection} showed that deep
denoising can \emph{destroy} task-relevant information even as conventional
image-quality metrics improve, and Kc \& Zeng~\cite{kc2024lgcho} found that
quarter-dose denoised CT can beat full-dose on PSNR and SSIM while remaining
\emph{inferior} on low-contrast detectability.  These works motivate the
hallucination-aware, detectability-based evaluation we report on this
benchmark's supervised and self-supervised models in
Section~\ref{sec:detect}.

\paragraph{Evaluation metrics.}
PSNR and SSIM are ubiquitous but favour pixel-level fidelity, potentially
penalising generative models that hallucinate plausible-but-incorrect
high-frequency detail.  Perceptual metrics such as
LPIPS~\cite{zhang2018unreasonable} and the noise power spectrum
(NPS)~\cite{solomon2012characterization} are more appropriate for
evaluating perceptual quality and texture authenticity, respectively.

% -----------------------------------------------------------------------
\section{Methods}

\subsection{Data and Preprocessing}

We use the TCIA LDCT-and-Projection-Data
dataset~\cite{moen2021low}, which contains paired low-dose and full-dose
reconstructions.  DICOM slices are read with \texttt{pydicom}, sorted by
$z$-position, and converted to Hounsfield units (HU).  HU values are
windowed to $[0, 1]$ via

\begin{equation}
  x = \operatorname{clamp}\!\left(\frac{u + b}{s},\; 0,\; 1\right),
  \label{eq:hu-norm}
\end{equation}

where $u$ is the raw HU value and $(b,s)$ is a fixed anatomy window preset.
The processed arrays are cached as an HDF5 file (one dataset per
patient/dose level) for efficient random access during training.

Patients are split at the patient level (not the slice level) into
training and validation sets using a configurable fraction
($\text{val\_fraction}=0.2$ by default) and a fixed random seed, preventing
any slice from a validation patient appearing in training.  During
training, $64\times64$ patches are sampled uniformly at random from each
slice.  During validation, full CT slices of variable size are processed
via the overlapped-tiling inference strategy described in
Section~\ref{sec:inference}.

\subsection{Architectures}

\paragraph{RED-CNN.}
Chen et al.'s~\cite{chen2017lowdose} Residual Encoder-Decoder CNN uses
$5\times5$ convolutions throughout.  The encoder consists of five
convolutional layers (96 filters each), and the decoder mirrors this
with five transposed-convolutional layers.  Skip connections are added at
three depths, and all activations use ReLU.

\paragraph{DnCNN.}
Zhang et al.'s~\cite{zhang2017beyond} denoising CNN uses 17 layers of
$3\times3$ convolutions.  The first layer has no batch normalisation;
intermediate layers each apply Conv-BN-ReLU; the final layer has neither.
The network predicts the noise component $n$ and returns
$\hat{x} = x - n$, a residual-learning formulation.

\paragraph{CTformer.}
Wang et al.'s~\cite{wang2022ctformer} architecture tokenises the input
with a strided convolution, applies a symmetric
encoder-bottleneck-decoder of Transformer blocks interleaved with
Token2Token Dilation (T2TD) modules, then detokenises with a transposed
convolution.  Each T2TD block applies a cyclic spatial shift and a dilated
re-tokenisation before multi-head self-attention.

\paragraph{U-Net.}
Our U-Net baseline uses three encoder stages, a two-layer bottleneck, and a
symmetric decoder with bilinear upsampling and concatenated skip
connections.  It is included for ablation: it shares the overall shape of
the velocity network inside our CFM model, so comparing U-Net (MSE) versus
CFM (flow-matching) isolates the contribution of the generative training
objective.

\paragraph{Conditional Flow Matching (CFM).}
\label{sec:cfm}
Conditional flow matching~\cite{lipman2022flow} frames denoising as
learning a time-dependent vector field $v_\theta$ that transports a
source distribution $p_0$ (LDCT images) to a target distribution $p_1$
(full-dose images).  The \emph{rectified flow}
formulation~\cite{liu2022flow} uses straight-line paths, so the
conditional vector field is simply

\begin{equation}
  v^* = x_1 - x_0,
  \label{eq:rectified-target}
\end{equation}

constant along the path.  The training objective is

\begin{equation}
  \mathcal{L}_\text{CFM}(\theta)
  = \mathbb{E}_{t, x_0, x_1}\bigl[\|v_\theta(x_t,\, x_0,\, t) - (x_1 - x_0)\|^2\bigr],
  \label{eq:cfm-loss}
\end{equation}

where $t \sim \mathcal{U}[0,1]$, the interpolant is
$x_t = (1-t)x_0 + t\,x_1$, and $x_0$ (the noisy LDCT patch) is passed as a
conditioning channel: the velocity network input is the two-channel
concatenation $[x_t; x_0]$.  At inference we integrate the learned ODE with
$K$ forward-Euler steps initialised at the noisy input.  This conditional
formulation requires paired clean targets and is the supervised baseline we
contrast against in Section~\ref{sec:method}.

\subsection{Training}

All models use the Adam optimiser at learning rate $10^{-4}$ and batch
size 4 (patches of $64\times64$).  Deterministic models (RED-CNN, DnCNN,
CTformer, U-Net) minimise pixel-level MSE; CFM uses
$\mathcal{L}_\text{CFM}$ (Eq.~\eqref{eq:cfm-loss}).

\paragraph{Supervision regimes.}
\emph{Supervised} training uses paired full-dose targets.
\emph{Noise2Void}~\cite{krull2019noise2void} replaces a fraction
$\rho=0.02$ of pixels per patch with a random neighbour within radius
$r=2$ and computes the loss against the original noisy value at those
blind-spot locations only.  \emph{Noise2Sim}~\cite{niu2020noise2sim}
constructs a self-supervised target from the centre pixels of non-local
self-similar patches.  \emph{ZS-N2N}~\cite{mansour2023zsn2n} is
architecture-agnostic: for each image a fresh three-layer residual network
($21.3\mathrm{K}$ parameters) is optimised for 2000 iterations on a
diagonally-downsampled Noise2Noise pair, then discarded.
\emph{Filter2Noise}~\cite{sun2025filter2noise} similarly optimises a small
single-image filter at test time.  Because CFM requires paired clean
targets, its blind-spot and similarity variants are undefined and omitted.

\subsection{Inference}
\label{sec:inference}

Full CT slices vary in size and are typically larger than the training
patch size.  We use an overlapped-tiling strategy to eliminate
grid-boundary artefacts: each patch of size $P\times P$ overlaps its
neighbours by a margin $m = P/4$, and predictions in the overlap region are
blended by a weighted mean over patch contributions.  Patches at the image
boundary retain their full extent: the margin is discarded only on sides
shared with a neighbouring patch, never on a side that abuts the image
edge.  An earlier version that discarded the margin on \emph{all} sides
left each slice's outer $m$-pixel border unpopulated (set to zero); because
the identity baseline does not pass through this routine, that artefact
biased the reconstruction metrics of every trained model.  All results
below use the corrected tiling.

\subsection{Evaluation Metrics}

\paragraph{PSNR / SSIM / RMSE.}
PSNR uses data range $R=1$ after normalisation; SSIM~\cite{wang2004image}
uses a Gaussian window ($\sigma=1.5$, window $11$); RMSE is
$\sqrt{\text{MSE}}$.

\paragraph{Gradient Magnitude Similarity Deviation (GMSD).}
GMSD~\cite{xue2014gradient} computes gradient magnitudes of the predicted
and reference images with Prewitt kernels and reports the standard
deviation of the gradient-magnitude-similarity map; lower is better and no
pretrained features are required.

\paragraph{Residual Spectral Ratio (RSR).}
\begin{equation}
  \text{RSR} = \frac{\mathbb{E}\bigl[|\mathcal{F}(\hat{x}-x_1)|^2\bigr]}
                    {\mathbb{E}\bigl[|\mathcal{F}(x_1)|^2\bigr] + \varepsilon},
  \label{eq:nps}
\end{equation}
where $\mathcal{F}$ is the 2-D DFT.  This measures the fraction of
clean-image spectral energy present in the residual error and is sensitive
to hallucinated high-frequency texture~\cite{gao2023corediff} that PSNR and
SSIM may miss.  RSR is a frequency-domain \emph{error} ratio against the
reference and should not be confused with the conventional CT noise-power
spectrum (NPS), which characterises the spectrum of the noise itself,
estimated from uniform-region ROIs; the latter is what we report as the NPS
radial centroid in the task-based evaluation of Section~\ref{sec:detect}.

% -----------------------------------------------------------------------
\section{Experiments}

\subsection{Setup}

Experiments use the TCIA LDCT-and-Projection-Data dataset~\cite{moen2021low}
(abdomen window), with a patient-level split (seed 0): $6{,}683$ training
and $1{,}476$ validation slices.  Trained models run for 50 epochs with
patch size 64, batch size 4, and Adam at $10^{-4}$; the CFM model
integrates the ODE with $K=20$ Euler steps during training and 5 at
evaluation.  The per-image methods (ZS-N2N, Filter2Noise) optimise one
network per validation image and are model-agnostic.  Validation metrics
are computed over full slices using the corrected overlapped inference of
Section~\ref{sec:inference} with margin $P/4 = 16$.  We report the gain
($\Delta$) of each metric relative to the unprocessed \emph{LDCT input},
the floor every method must beat.

\subsection{Results}
\label{sec:results}

\begin{table*}[t]
  \centering
  \small
  \caption{Validation results on TCIA LDCT (abdomen; patient-level split,
           seed 0; 50 epochs), grouped by supervision regime.  Means over
           the validation set on full slices via overlapped tiling
           (corrected border handling).  PSNR shows the gain over the
           \emph{LDCT input} floor in parentheses.  $\uparrow$/$\downarrow$:
           higher/lower is better.  \textbf{Bold}: best denoiser per metric.
           ZS-N2N and Filter2Noise are per-image and model-agnostic (single
           row each).  $\dagger$: run truncated by a crash, reported at its
           last logged epoch (CTformer/Supervised ep.\,37,
           CTformer/Noise2Sim ep.\,26).  $\ddagger$: CFM crashed at ep.\,15
           and had not converged; shown for reference.  CFM requires paired
           targets, so it has no blind-spot or similarity variant.  SSFlow
           (ours) rows use one-step inference at 50 epochs; the one- vs.\
           multi-step ablation is in Section~\ref{sec:method}.
           \underline{Underline}: best \emph{label-free} result.}
  \label{tab:results}
  \begin{tabular}{llrlcccc}
    \toprule
    \textbf{Model} & \textbf{Supervision} & \textbf{Params} &
    PSNR\,$\uparrow$ (\,$\Delta$) & SSIM\,$\uparrow$ & RMSE\,$\downarrow$ &
    GMSD\,$\downarrow$ & RSR\,$\downarrow$ \\
    \midrule
    LDCT input & --- (no denoising) & --- & 31.44 & 0.893 & 0.0274 & 0.166 & 0.0079 \\
    \midrule
    RED-CNN~\cite{chen2017lowdose}   & Supervised & $1.85\mathrm{M}$ & \textbf{35.48} \,(+4.04) & \textbf{0.940} & 0.0171 & 0.166 & 0.0031 \\
    U-Net                            & Supervised & $1.95\mathrm{M}$ & 35.47 \,(+4.04) & 0.940 & \textbf{0.0171} & 0.178 & \textbf{0.0030} \\
    DnCNN~\cite{zhang2017beyond}     & Supervised & $0.56\mathrm{M}$ & 35.40 \,(+3.97) & 0.938 & 0.0172 & 0.184 & 0.0031 \\
    CTformer~\cite{wang2022ctformer}$^{\dagger}$ & Supervised & $0.40\mathrm{M}$ & 34.78 \,(+3.34) & 0.917 & 0.0184 & 0.263 & 0.0036 \\
    CFM~\cite{lipman2022flow}$^{\ddagger}$       & Supervised & $2.11\mathrm{M}$ & 33.87 \,(+2.43) & 0.923 & 0.0206 & 0.248 & 0.0044 \\
    \midrule
    U-Net                            & Noise2Sim~\cite{niu2020noise2sim} & $1.95\mathrm{M}$ & 32.61 \,(+1.18) & 0.910 & 0.0238 & 0.169 & 0.0060 \\
    DnCNN~\cite{zhang2017beyond}     & Noise2Sim~\cite{niu2020noise2sim} & $0.56\mathrm{M}$ & 32.14 \,(+0.70) & 0.891 & 0.0251 & 0.178 & 0.0067 \\
    RED-CNN~\cite{chen2017lowdose}   & Noise2Sim~\cite{niu2020noise2sim} & $1.85\mathrm{M}$ & 31.94 \,(+0.51) & 0.904 & 0.0257 & \textbf{0.162} & 0.0070 \\
    CTformer~\cite{wang2022ctformer}$^{\dagger}$ & Noise2Sim~\cite{niu2020noise2sim} & $0.40\mathrm{M}$ & 31.09 \,($-0.35$) & 0.885 & 0.0282 & 0.247 & 0.0083 \\
    \midrule
    SSFlow $r{=}2$ (ours) & Self-sup.\ flow & $2.11\mathrm{M}$ & \underline{32.83} \,(+1.39) & \underline{0.917} & \underline{0.0231} & 0.186 & \underline{0.0057} \\
    SSFlow $r{=}3$ (ours) & Self-sup.\ flow & $2.11\mathrm{M}$ & 32.28 \,(+0.85) & 0.914 & 0.0246 & 0.189 & 0.0064 \\
    SSFlow $r{=}1$ (ours) & Self-sup.\ flow & $2.11\mathrm{M}$ & 32.20 \,(+0.77) & 0.907 & 0.0249 & 0.183 & 0.0066 \\
    \midrule
    RED-CNN~\cite{chen2017lowdose}   & Noise2Void~\cite{krull2019noise2void} & $1.85\mathrm{M}$ & 31.52 \,(+0.08) & 0.895 & 0.0271 & 0.175 & 0.0077 \\
    U-Net                            & Noise2Void~\cite{krull2019noise2void} & $1.95\mathrm{M}$ & 31.49 \,(+0.05) & 0.893 & 0.0271 & 0.188 & 0.0078 \\
    CTformer~\cite{wang2022ctformer} & Noise2Void~\cite{krull2019noise2void} & $0.40\mathrm{M}$ & 31.28 \,($-0.15$) & 0.894 & 0.0277 & 0.223 & 0.0081 \\
    DnCNN~\cite{zhang2017beyond}     & Noise2Void~\cite{krull2019noise2void} & $0.56\mathrm{M}$ & 30.98 \,($-0.46$) & 0.885 & 0.0288 & 0.198 & 0.0087 \\
    \midrule
    Filter2Noise~\cite{sun2025filter2noise} & Zero-shot & --- & 31.58 \,(+0.14) & 0.896 & 0.0269 & 0.176 & 0.0076 \\
    ZS-N2N~\cite{mansour2023zsn2n}          & Zero-shot & $21.3\mathrm{K}$ & 31.17 \,($-0.27$) & 0.892 & 0.0281 & 0.182 & 0.0083 \\
    \bottomrule
  \end{tabular}
\end{table*}

Table~\ref{tab:results} yields four findings.  \textbf{First}, supervised
denoising gains $\sim$4\,dB PSNR and $+0.046$ SSIM over the noisy input, and
RED-CNN, U-Net, and DnCNN finish within $0.1$\,dB of one another despite a
$3\times$ range in parameter count, so the objective dominates the backbone.
\textbf{Second}, among label-free methods only the correlated-noise-aware
Noise2Sim consistently exceeds the noisy floor ($+0.5$ to $+1.2$\,dB),
whereas blind-spot Noise2Void is flat-to-negative ($-0.46$\,dB and $-0.008$
SSIM on DnCNN) and the zero-shot ZS-N2N sits essentially at the floor; this
is direct evidence that FBP-correlated CT noise violates the
pixel-independence assumption these methods rely on.  \textbf{Third}, the
$3$--$4$\,dB gap between the best self-supervised result and the supervised
upper bound quantifies the headroom for correlated-noise-aware
self-supervision, which motivates the method of Section~\ref{sec:method}.
\textbf{Fourth}, that method---our self-supervised rectified flow
(SSFlow)---closes part of this gap: at $r{=}2$ it reaches $+1.39$\,dB, the best
label-free result in the table and ahead of the strongest Noise2Sim backbone
($+1.18$\,dB), while also leading every label-free method on SSIM, RMSE, and
RSR.
CTformer and CFM runs were truncated by crashes and are reported at their
last logged epoch; completing them, together with multi-anatomy and
multi-seed evaluation, is in progress.

\paragraph{Cross-validation robustness.}
To check that the single-split ordering above is not an artefact of one
patient partition, we repeat the regime comparison under a three-fold,
patient-level cross-validation at a reduced 15-epoch budget
(Table~\ref{tab:cv}).  The folds differ in difficulty---their unprocessed
LDCT-input PSNR is $31.44$, $30.04$, and $29.27$\,dB---so we report the gain
over each fold's \emph{own} input floor, averaged across folds ($\pm$ s.d.).
The DnCNN and RED-CNN backbones, which completed all three folds, preserve the
regime ordering of Table~\ref{tab:results} with a tight per-fold spread:
supervised training gains $+4.0$ to $+4.2$\,dB ($\pm0.22$); Noise2Sim is again
the only label-free method to clear the floor by a decisive margin
($+1.74\pm0.09$\,dB on RED-CNN); and Noise2Void, Filter2Noise, and ZS-N2N stay
within noise of the input ($|\Delta\text{PSNR}|\le0.32$\,dB).  Absolute PSNR
runs $\sim$1\,dB below Table~\ref{tab:results} because these models train for 15
rather than 50 epochs, but the \emph{relative} ordering---the paper's central
claim---is unchanged across every fold.

\begin{table*}[t]
  \centering
  \small
  \caption{Three-fold patient-level cross-validation on TCIA LDCT (abdomen;
           15 epochs).  $\Delta$PSNR and SSIM are the mean ($\pm$ s.d.) across
           the three folds; the PSNR gain is measured against each fold's own
           LDCT-input floor ($31.44/30.04/29.27$\,dB for folds $1/2/3$).  PSNR,
           RMSE, and RSR are fold means.  Only DnCNN and RED-CNN completed
           all three folds and are shown; U-Net had incomplete fold coverage at
           export.  Filter2Noise and ZS-N2N are per-image and model-agnostic
           (single row each).  \textbf{Bold}: best per column.
           \underline{Underline}: best \emph{label-free} PSNR gain.}
  \label{tab:cv}
  \begin{tabular}{llccccc}
    \toprule
    \textbf{Model} & \textbf{Supervision} &
    $\Delta$PSNR\,$\uparrow$ (dB) & PSNR\,$\uparrow$ & SSIM\,$\uparrow$ &
    RMSE\,$\downarrow$ & RSR\,$\downarrow$ \\
    \midrule
    RED-CNN~\cite{chen2017lowdose} & Supervised & \textbf{+4.19 $\pm$ 0.22} & \textbf{34.43} & \textbf{0.927 $\pm$ 0.013} & \textbf{0.0195} & \textbf{0.0042} \\
    DnCNN~\cite{zhang2017beyond}   & Supervised & +4.04 $\pm$ 0.22 & 34.29 & 0.924 $\pm$ 0.015 & 0.0199 & 0.0044 \\
    \midrule
    RED-CNN~\cite{chen2017lowdose} & Noise2Sim~\cite{niu2020noise2sim} & \underline{+1.74 $\pm$ 0.09} & 31.99 & 0.900 $\pm$ 0.015 & 0.0261 & 0.0077 \\
    DnCNN~\cite{zhang2017beyond}   & Noise2Sim~\cite{niu2020noise2sim} & +1.42 $\pm$ 0.28 & 31.66 & 0.895 $\pm$ 0.015 & 0.0271 & 0.0082 \\
    \midrule
    RED-CNN~\cite{chen2017lowdose} & Noise2Void~\cite{krull2019noise2void} & +0.32 $\pm$ 0.10 & 30.57 & 0.883 $\pm$ 0.019 & 0.0306 & 0.0105 \\
    DnCNN~\cite{zhang2017beyond}   & Noise2Void~\cite{krull2019noise2void} & $-0.02$ $\pm$ 0.06 & 30.23 & 0.875 $\pm$ 0.018 & 0.0320 & 0.0114 \\
    \midrule
    Filter2Noise~\cite{sun2025filter2noise} & Zero-shot & +0.26 $\pm$ 0.11 & 30.51 & 0.881 $\pm$ 0.016 & 0.0309 & 0.0105 \\
    ZS-N2N~\cite{mansour2023zsn2n}          & Zero-shot & $-0.09$ $\pm$ 0.20 & 30.16 & 0.879 $\pm$ 0.015 & 0.0319 & 0.0112 \\
    \bottomrule
  \end{tabular}
\end{table*}

\subsection{Task-based detectability}
\label{sec:detect}

Pixel-fidelity gains need not be diagnostic gains.  We test this directly with
a Channelized Hotelling Observer (CHO) on a signal-known-exactly /
background-known-statistically (SKE/BKS) task.  No CT forward projector is
needed: because the dataset provides real paired low/full acquisitions, the
real spatially-correlated FBP noise is $n = x_\text{low} - x_\text{full}$, so a
signal-present low-dose image is \emph{literally} $x_\text{low} + s$ for a known
lesion $s$ (the same correlated noise is re-attached by construction), and the
signal-absent image is the untouched acquisition.  We insert Gaussian lesions of
radius $4$\,px at $40/80/160$\,HU on flat in-body sites ($6$ per slice, $200$
slices per run), extract present/absent ROIs at the \emph{input}, \emph{denoised}
and \emph{clean} stages, channelise with ten Laguerre-Gauss channels, and compute
the detectability index $d'$.  The observer is validated against the closed-form
white-noise value $d' = \lVert s\rVert / \sigma$ as a unit test before use.  We
run it on the completed three-seed benchmark models of
Table~\ref{tab:sweep-benchmark} (the three CNN architectures) and report the
mean over architectures and seeds; CTformer reproduces the same erosion and is
held out only to keep the aggregate architecturally homogeneous.  The
per-image methods (ZS-N2N, Filter2Noise) are omitted---each CHO realisation would
retrain their per-image network thousands of times---and the generative
SSFlow/CFM detectability, which can \emph{fabricate} as well as erode, is reported
separately.

\begin{table*}[t]
  \centering
  \small
  \caption{Task-based detectability on the completed benchmark models
           (Table~\ref{tab:sweep-benchmark}); mean over $3$ CNN architectures
           $\times$ $3$ seeds ($27$ runs; CTformer, now complete, reproduces the
           same erosion and is held out to keep the aggregate CNN-only).
           \emph{Top:} CHO detectability index $d'$ for the noisy
           input (floor) and clean reference (ceiling) at three lesion
           contrasts; these are method-independent and carry a large
           between-split (patient-cohort) s.d.\ ($\pm1.3$ at $160$\,HU), which is
           why we report the ratio below.  \emph{Bottom:} detectability
           \emph{preserved}, $d'_\text{denoised}/d'_\text{input}$
           (mean\,$\pm$\,s.d.; $1.0=$ unchanged, $<1=$ eroded); the denoised
           noise-power-spectrum radial centroid (input $0.112$; lower $=$ more
           low-frequency ``waxy'' texture); and the \emph{fabrication} $d'$ from
           a CHO run on \emph{signal-absent} ROIs (input floor $0.028$; $>$floor
           $=$ invented structure).  Architecture differences are within s.d.,
           so rows are aggregated by supervision regime.  $d'_\text{denoised}$ is
           below $d'_\text{input}$ in all $27$ runs and never exceeds the clean
           ceiling, while fabrication $d'$ never rises above the floor.}
  \label{tab:detect}
  \begin{tabular}{lccccc}
    \toprule
     & \multicolumn{3}{c}{lesion contrast (HU)} & NPS & fab. \\
    \cmidrule(lr){2-4}
     & $40$ & $80$ & $160$ & centroid\,$\downarrow$ & $d'\,\downarrow$ \\
    \midrule
    \multicolumn{6}{l}{\emph{Reference $d'$ (method-independent)}} \\
    \quad input (floor)   & $1.51$ & $3.03$ & $6.06$ & $0.112$ & $0.028$ \\
    \quad clean (ceiling) & $1.54$ & $3.09$ & $6.18$ & --- & --- \\
    \midrule
    \multicolumn{6}{l}{\emph{Detectability preserved $d'_\text{den}/d'_\text{in}$}} \\
    \quad Supervised & $1.024\,{\pm}\,0.008$ & $1.016\,{\pm}\,0.007$ & $0.934\,{\pm}\,0.033$ & $0.087\,{\pm}\,0.004$ & $0.022\,{\pm}\,0.003$ \\
    \quad Noise2Sim  & $1.008\,{\pm}\,0.006$ & $0.995\,{\pm}\,0.006$ & $0.919\,{\pm}\,0.030$ & $0.100\,{\pm}\,0.005$ & $0.028\,{\pm}\,0.006$ \\
    \quad Noise2Void & $1.003\,{\pm}\,0.006$ & $0.993\,{\pm}\,0.007$ & $0.919\,{\pm}\,0.030$ & $0.109\,{\pm}\,0.006$ & $0.033\,{\pm}\,0.006$ \\
    \bottomrule
  \end{tabular}
\end{table*}

Table~\ref{tab:detect} shows four things.  \textbf{First}, the detectability
\emph{headroom} is small: the clean reference exceeds the noisy input by only
$\sim$2\% ($d' = 6.18$ vs.\ $6.06$ at $160$\,HU), so at this dose the inserted
lesions are nearly as detectable in the noisy image as in the noise-free one.
\textbf{Second}, although these same models gain $4.0$--$4.3$\,dB PSNR
(supervised) and $1.7$--$1.8$\,dB (Noise2Sim) over the input
(Table~\ref{tab:sweep-benchmark}), none improves detectability at the
supra-threshold operating point: $d'_\text{denoised} < d'_\text{input}$ in all
$27$ runs---a $7$--$8\%$ loss at $160$\,HU---and none exceeds the clean ceiling,
while the fabrication $d'$ on signal-absent ROIs stays at its floor
($0.022$--$0.033$ vs.\ the $0.028$ input floor), so the deterministic denoisers
\emph{erode} detectability without \emph{fabricating} it.  The erosion is
contrast-dependent: near the detection threshold ($40$\,HU) detectability is
preserved to within $1$--$2\%$, so the loss concentrates precisely at the
high-contrast operating point where PSNR and SSIM report their largest, most
confident gains.  \textbf{Third}, the supervised models that win most on PSNR
also shift the noise-power-spectrum centroid furthest toward low frequency
($0.112 \!\to\! 0.087$, a $22\%$ drop, vs.\ $0.112 \!\to\! 0.109$ for
Noise2Void), the blotchy / ``waxy'' texture radiologists distrust.
\textbf{Fourth}, the differences \emph{between} supervision regimes are not
statistically resolved at this sample size: the preserved-detectability ratios
at $160$\,HU ($0.934\,{\pm}\,0.033$ supervised, $0.919\,{\pm}\,0.030$
Noise2Sim, $0.919\,{\pm}\,0.030$ Noise2Void) overlap well within one standard
deviation, so the data support the \emph{erosion} claim---every regime sits
below $1.0$---but not a fine-grained regime-vs-regime detectability ranking;
only the NPS-centroid shift cleanly separates supervised from blind-spot
training.  Reporting the \emph{preserved} ratio
$d'_\text{denoised}/d'_\text{input}$ rather than the raw index matters here: it
cancels the large between-split variation in absolute $d'$ (s.d.\ $\pm1.31$ at
$160$\,HU) and is stable across seeds ($\pm0.003$--$0.033$), isolating the
denoiser's effect.  None of this contrast-dependent erosion or texture shift
appears in PSNR or SSIM, which rise monotonically; task-based $d'$ is therefore
a necessary complement to them, not an optional one.

% -----------------------------------------------------------------------
\section{An Unconditional Rectified Flow is a One-Step Noise2Noise Denoiser}
\label{sec:method}

The results above leave a $3$--$4$\,dB gap between the best label-free
method and supervised training.  Our methodological contribution is an
\emph{observation}: the velocity field of an \emph{unconditional} rectified
flow, trained on two noisy observations of the same signal, has a minimiser
that recovers the Noise2Noise posterior mean in a \emph{single} step
(Section~\ref{sec:rf-n2n})---no clean targets, and no ODE integration
required.  We then instantiate this observation as SSFlow, a denoiser
designed for the correlated noise of FBP-reconstructed LDCT
(Section~\ref{sec:ss-pairs}): it pairs the unconditional flow with a
\emph{similarity-based} construction of the noisy pairs that keeps the
paired noise independent despite spatial correlation.  An ablation
(Section~\ref{sec:ss-pairs}) confirms the consequence of the observation
that one might not expect of a ``flow''---multi-step refinement does not
help---so SSFlow is best read as the empirical realisation of the one-step
result rather than as iterative sampling.

\subsection{Rectified flow as a Noise2Noise estimator}
\label{sec:rf-n2n}
Let $s$ be the (unobserved) clean slice and let $x_0=s+n_0,\ x_1=s+n_1$ be
two noisy observations with $\mathbb{E}[n_i\mid s]=0$ and
$n_0\perp n_1\mid s$.  Using the same straight-line interpolant
$z_t=(1-t)x_0+t\,x_1$ as Eq.~\eqref{eq:rectified-target}, we train an
\emph{unconditional} velocity network $v_\theta(z,t)$ (inputs: the
interpolant $z$ and time $t$ only, in contrast to the conditional field of
Eq.~\eqref{eq:cfm-loss}) with

\begin{equation}
  \mathcal{L}(\theta)=
  \mathbb{E}_{s,n_0,n_1,\,t\sim\mathcal{U}[0,1]}
  \big\|\,v_\theta(z_t,t)-(x_1-x_0)\,\big\|^2 .
  \label{eq:ss-loss}
\end{equation}

\paragraph{Proposition.}
The minimiser of \eqref{eq:ss-loss} is the conditional expectation
$v^\star(z,t)=\mathbb{E}[\,x_1-x_0\mid z_t=z\,]$.  At $t=0$, $z_0=x_0$, so

\begin{equation}
  x_0 + v^\star(x_0,0)
  = \mathbb{E}[x_1\mid x_0]
  = \mathbb{E}[s\mid x_0],
  \label{eq:ss-mmse}
\end{equation}

the MMSE denoiser.  The last equality is the key step: by the tower property,
$\mathbb{E}[x_1\mid x_0]=\mathbb{E}\!\left[\mathbb{E}[x_1\mid s,x_0]\mid
x_0\right]=\mathbb{E}\!\left[s+\mathbb{E}[n_1\mid s,x_0]\mid x_0\right]
=\mathbb{E}[s\mid x_0]$, where $\mathbb{E}[n_1\mid s,x_0]=\mathbb{E}[n_1\mid
s]=0$ because $n_1\perp n_0\mid s$ (so conditioning on $x_0=s+n_0$ adds
nothing beyond $s$) and $\mathbb{E}[n_i\mid s]=0$.  A single velocity
evaluation thus recovers the Noise2Noise~\cite{lehtinen2018noise2noise}
optimum from noisy pairs alone.

\paragraph{Why the flow must be unconditional.}
If the network is additionally given $x_0$ (as in
Eq.~\eqref{eq:cfm-loss}), then $(z_t,x_0,t)$ determines the target exactly,
$x_1=(z_t-(1-t)x_0)/t$: the regression target is fully observed and the
network reproduces the \emph{noisy} $x_1$ rather than averaging over it,
destroying \eqref{eq:ss-mmse}.  Dropping the conditioning makes each $z_t$
consistent with many pairs, forcing the network to predict their average
velocity, where the independent noise cancels.

\subsection{Similarity-based pairs and inference}
\label{sec:ss-pairs}
Equation~\eqref{eq:ss-mmse} requires two observations whose noise is
independent given the signal.  FBP smears each detector reading across
neighbouring pixels, so spatial-neighbour pairs
(Neighbor2Neighbor~\cite{huang2021neighbor2neighbor},
ZS-N2N~\cite{mansour2023zsn2n}) are only valid when the noise correlation
length is below the pixel pitch.  We instead adopt the non-local
construction of Noise2Sim~\cite{niu2020noise2sim}: for each query patch we
retrieve its $K$ most similar patches elsewhere in the slice, \emph{excluding}
a radius $r$ around the query so matched patches lie beyond the noise
correlation length.  Self-similarity matches the signal while spatial
separation decorrelates the noise, yielding a valid pair $(x_0,x_1)$ from a
single image; $r$ is the key correlated-noise hyper-parameter.  At
inference, the one-step estimate is $\hat s = x + v_\theta(x,0)$, optionally
refined with a few Euler steps of the learned ODE initialised at the noisy
input; we do \emph{not} integrate to the noisy endpoint distribution.

\paragraph{Ablation.}
By \eqref{eq:ss-mmse}, the one-step estimate coincides with
similarity-based Noise2Noise regression; the flow is justified only if
\emph{multi-step refinement improves the structured residual} that one-step
MMSE leaves behind.  We isolate this on a $3{\times}2$ grid over the exclusion
radius $r\in\{1,2,3\}$ and the number of inference steps $\{1,4\}$
(Table~\ref{tab:ssflow-steps}), against
direct Noise2Sim and the supervised upper bound.  Two findings emerge.
\emph{(i) The decorrelated pairing is the active ingredient.}  PSNR follows an
inverted-U in $r$, peaking at $r{=}2$ ($+1.39$\,dB, Table~\ref{tab:ssflow-steps});
$r{=}1$---which excludes only the query pixel, recovering the Noise2Sim
target---is $0.6$\,dB worse, confirming that matching \emph{beyond} the FBP
correlation length is what helps.  \emph{(ii) Multi-step refinement does not
help; it hurts.}  Four-step Euler inference is uniformly below one-step at
matched training and \emph{degrades} as training proceeds, falling to $+0.52$,
$+0.48$, and $+0.08$\,dB at $r{=}1,2,3$ (four-step runs ongoing, ep.\,39--48)
versus $+0.77$, $+1.39$, $+0.85$ for one-step (ep.\,50).  The one-step
posterior mean of \eqref{eq:ss-mmse}---not iterative sampling---is therefore
the operative estimator: a null multi-step effect that narrows the
contribution to the correlated-noise-aware pairing, in line with
``no-new-denoiser'' reports that stochastic sampling trades fidelity for
texture.  Consistent with that trade, val PSNR peaks early and then declines
while GMSD keeps improving, so we checkpoint and report the
best-validation-PSNR epoch.

\subsection{Per-sweep result tables}
\label{sec:sweep-tables}
The four grid sweeps that drive this study each populate a dedicated table,
reported here as their runs complete.  Table~\ref{tab:ssflow-steps} is the
inference-steps ablation above (\texttt{sweep\_hallucination.yml}), filled from
the current export with an explicit \emph{epoch} column because the four-step
arm is still training; Table~\ref{tab:flow-vs-reg}
(\texttt{sweep\_flow\_vs\_reg.yml}) is its regression control, holding the
similarity pairing fixed and swapping the flow estimator for a direct Noise2Sim
regressor to isolate \emph{why} SSFlow trails Noise2Sim;
Table~\ref{tab:sweep-benchmark} (\texttt{sweep.yml}) and
Table~\ref{tab:sweep-periimage} (\texttt{sweep\_periimage.yml}) give the
architecture\,$\times$\,supervision and per-image grids as scaffolds---rows and
columns fixed, cells (``---'') populated as those sweeps finish.

\begin{table*}[t]
  \centering
  \small
  \caption{\textbf{SSFlow inference-steps ablation} (\texttt{sweep\_hallucination.yml}):
           one-step posterior mean vs.\ multi-step Euler refinement across the
           decorrelation radius $r$, on TCIA LDCT (abdomen; seed 0).  \emph{In
           progress}---the \textbf{Epoch} column reports each run's last logged
           epoch: the one-step runs have reached $50$, the four-step arm is
           still training ($39$--$48$), so the two blocks are not yet at matched
           budget.  PSNR shows the gain over the \emph{LDCT input} floor in
           parentheses.  The fabrication $d'$ column (generative hallucination
           index from the \texttt{--eval-steps-sweep} detectability pass) is
           pending in this export.  \textbf{Bold}: best one-step row per metric.
           $\uparrow$/$\downarrow$: higher/lower is better.}
  \label{tab:ssflow-steps}
  \begin{tabular}{cclrccccc}
    \toprule
    \textbf{Steps} & \textbf{Radius $r$} & \textbf{Epoch} &
    PSNR\,$\uparrow$ (\,$\Delta$) & SSIM\,$\uparrow$ & RMSE\,$\downarrow$ &
    GMSD\,$\downarrow$ & RSR\,$\downarrow$ & fab.\ $d'$\,$\downarrow$ \\
    \midrule
    \multicolumn{9}{l}{\emph{Reference}} \\
    \quad LDCT input & --- & --- & 31.44 & 0.893 & 0.0274 & 0.166 & 0.0079 & --- \\
    \midrule
    \multicolumn{9}{l}{\emph{One-step posterior mean}} \\
    $1$ & $1$ & $50$ & 32.20 \,(+0.77) & 0.907 & 0.0249 & 0.183 & 0.0066 & --- \\
    $1$ & $2$ & $50$ & \textbf{32.83} \,(+1.39) & \textbf{0.917} & \textbf{0.0231} & 0.186 & \textbf{0.0057} & --- \\
    $1$ & $3$ & $50$ & 32.28 \,(+0.85) & 0.914 & 0.0246 & 0.189 & 0.0064 & --- \\
    \midrule
    \multicolumn{9}{l}{\emph{Four-step Euler (still training)}} \\
    $4$ & $1$ & $39$ & 31.96 \,(+0.52) & 0.903 & 0.0257 & 0.176 & 0.0070 & --- \\
    $4$ & $2$ & $44$ & 31.92 \,(+0.48) & 0.902 & 0.0257 & 0.192 & 0.0070 & --- \\
    $4$ & $3$ & $48$ & 31.52 \,(+0.08) & 0.895 & 0.0268 & 0.195 & 0.0076 & --- \\
    \bottomrule
  \end{tabular}
\end{table*}

\begin{table*}[t]
  \centering
  \small
  \caption{\textbf{Flow-vs-regression control} (\texttt{sweep\_flow\_vs\_reg.yml}):
           \emph{why} does SSFlow trail Noise2Sim when the one-step theorem
           (Section~\ref{sec:rf-n2n}) says they share the MMSE target
           $\mathbb{E}[s\mid x]$?  We hold the similarity pairing fixed and vary
           only the estimator.  The \emph{regression} arm is a direct Noise2Sim
           regressor (RED-CNN backbone, $1.85\mathrm{M}$ params, capacity-comparable
           to SSFlow's $2.11\mathrm{M}$ velocity net) trained on SSFlow's exact
           decorrelated pairs; the \emph{flow} arm is one-step SSFlow
           (\texttt{sweep\_hallucination.yml}).  Both sweep the exclude-radius
           $r\in\{1,2,3\}$ over $3$ seeds ($9$ runs each), so the two arms overlay
           cell-for-cell.  \emph{Scaffold}---data cells (``---'') are populated as
           the sweeps finish; means will be mean\,$\pm$\,s.d.\ over $3$ seeds with
           PSNR gain over the $30.25$\,dB $3$-seed input floor.  The decisive read
           is regression$(r)$ vs.\ flow$(r)$ at matched $r$: if regression holds
           near its seed-0 anchor ($r{=}1$ reproduces standard Noise2Sim,
           $\sim{+}1.81$\,dB on RED-CNN) while flow stays at its one-step seed-0
           anchors ($+0.77/+1.39/+0.85$\,dB at $r{=}1/2/3$;
           Table~\ref{tab:ssflow-steps}), the \emph{flow objective} is the cost;
           if regression also falls at $r{=}2/3$, the \emph{decorrelated pairs}
           are.  The fabrication $d'$ (from \texttt{--eval-detectability}) checks
           that neither estimator invents structure.  $\dagger$: provisional
           $r{=}2$ flow value carried from the main sweep's 3-seed
           \texttt{flowmatching+n2sim} cell ($5$-step Euler eval, $2$ of $3$ seeds
           finished; fabrication $d'$ at the input floor); it is the only
           multi-seed SSFlow number to date but is \emph{multi-step}, so the
           dedicated one-step arm replaces it.  $\uparrow$/$\downarrow$:
           higher/lower is better.}
  \label{tab:flow-vs-reg}
  \begin{tabular}{lcrccccc}
    \toprule
    \textbf{Estimator} & \textbf{Radius $r$} &
    PSNR\,$\uparrow$ (\,$\Delta$) & SSIM\,$\uparrow$ & RMSE\,$\downarrow$ &
    GMSD\,$\downarrow$ & RSR\,$\downarrow$ & fab.\ $d'$\,$\downarrow$ \\
    \midrule
    \multicolumn{8}{l}{\emph{Reference}} \\
    \quad LDCT input & --- & 30.25 & 0.876 & 0.0320 & 0.174 & 0.0115 & 0.028 \\
    \midrule
    \multicolumn{8}{l}{\emph{Regression (Noise2Sim, RED-CNN) --- same pairs, direct MMSE}} \\
    Regression & $1$ & --- & --- & --- & --- & --- & --- \\
    Regression & $2$ & --- & --- & --- & --- & --- & --- \\
    Regression & $3$ & --- & --- & --- & --- & --- & --- \\
    \midrule
    \multicolumn{8}{l}{\emph{Flow (SSFlow) --- same pairs, rectified-flow objective; one-step unless marked}} \\
    Flow & $1$ & --- & --- & --- & --- & --- & --- \\
    Flow & $2^{\dagger}$ & 31.27 \,($+0.92{\pm}0.03$) & $0.889{\pm}0.025$ & $0.0287$ & $0.187$ & $0.0097$ & $0.030$ \\
    Flow & $3$ & --- & --- & --- & --- & --- & --- \\
    \bottomrule
  \end{tabular}
\end{table*}

\begin{table*}[t]
  \centering
  \small
  \caption{\textbf{Main benchmark sweep} (\texttt{sweep.yml}): architecture
           $\times$ supervision grid ($5$ models $\times$ $3$ modes $\times$ $3$
           seeds; CFM is paired-target only, so it has no self-supervised
           variant).  All five architectures are complete over $3$ seeds and
           reported as mean\,$\pm$\,s.d.  The
           PSNR column gives the absolute value with the gain over the LDCT
           input floor in parentheses.  Per-image methods are in
           Table~\ref{tab:sweep-periimage}.  \textbf{Bold}: best $\Delta$PSNR
           per supervision block.  $\uparrow$/$\downarrow$: higher/lower is
           better.}
  \label{tab:sweep-benchmark}
  \begin{tabular}{llrccccc}
    \toprule
    \textbf{Model} & \textbf{Supervision} & \textbf{Params} &
    PSNR\,$\uparrow$ (\,$\Delta$) & SSIM\,$\uparrow$ & RMSE\,$\downarrow$ &
    GMSD\,$\downarrow$ & RSR\,$\downarrow$ \\
    \midrule
    LDCT input & --- (no denoising) & --- & $30.25$ & $0.876$ & $0.0320$ & $0.174$ & $0.0115$ \\
    \midrule
    RED-CNN~\cite{chen2017lowdose}   & Supervised & $1.85\mathrm{M}$ & \textbf{34.50 \,($+4.25{\pm}0.23$)} & $0.927{\pm}0.013$ & $0.0194$ & $0.167$ & $0.0041$ \\
    U-Net                            & Supervised & $1.95\mathrm{M}$ & 34.26 \,($+4.02{\pm}0.22$) & $0.925{\pm}0.014$ & $0.0199$ & $0.188$ & $0.0044$ \\
    DnCNN~\cite{zhang2017beyond}     & Supervised & $0.56\mathrm{M}$ & 34.23 \,($+3.99{\pm}0.18$) & $0.924{\pm}0.014$ & $0.0200$ & $0.198$ & $0.0044$ \\
    CTformer~\cite{wang2022ctformer} & Supervised & $0.40\mathrm{M}$ & 32.01 \,($+1.77{\pm}0.35$) & $0.904{\pm}0.013$ & $0.0255$ & $0.249$ & $0.0070$ \\
    CFM~\cite{lipman2022flow}        & Supervised & $2.11\mathrm{M}$ & 33.09 \,($+2.84{\pm}0.22$) & $0.915{\pm}0.013$ & $0.0228$ & $0.192$ & $0.0057$ \\
    \midrule
    RED-CNN~\cite{chen2017lowdose}   & Noise2Sim~\cite{niu2020noise2sim} & $1.85\mathrm{M}$ & 32.05 \,($+1.81{\pm}0.10$) & $0.901{\pm}0.015$ & $0.0259$ & $0.163$ & $0.0075$ \\
    U-Net                            & Noise2Sim~\cite{niu2020noise2sim} & $1.95\mathrm{M}$ & \textbf{32.08 \,($+1.83{\pm}0.08$)} & $0.901{\pm}0.015$ & $0.0259$ & $0.174$ & $0.0075$ \\
    DnCNN~\cite{zhang2017beyond}     & Noise2Sim~\cite{niu2020noise2sim} & $0.56\mathrm{M}$ & 31.93 \,($+1.69{\pm}0.09$) & $0.898{\pm}0.015$ & $0.0263$ & $0.187$ & $0.0078$ \\
    CTformer~\cite{wang2022ctformer} & Noise2Sim~\cite{niu2020noise2sim} & $0.40\mathrm{M}$ & 31.04 \,($+0.79{\pm}0.13$) & $0.887{\pm}0.018$ & $0.0289$ & $0.232$ & $0.0092$ \\
    \midrule
    RED-CNN~\cite{chen2017lowdose}   & Noise2Void~\cite{krull2019noise2void} & $1.85\mathrm{M}$ & \textbf{30.67 \,($+0.42{\pm}0.06$)} & $0.883{\pm}0.018$ & $0.0303$ & $0.177$ & $0.0103$ \\
    U-Net                            & Noise2Void~\cite{krull2019noise2void} & $1.95\mathrm{M}$ & 30.54 \,($+0.29{\pm}0.04$) & $0.883{\pm}0.018$ & $0.0307$ & $0.198$ & $0.0104$ \\
    DnCNN~\cite{zhang2017beyond}     & Noise2Void~\cite{krull2019noise2void} & $0.56\mathrm{M}$ & 30.19 \,($-0.06{\pm}0.03$) & $0.875{\pm}0.018$ & $0.0321$ & $0.224$ & $0.0116$ \\
    CTformer~\cite{wang2022ctformer} & Noise2Void~\cite{krull2019noise2void} & $0.40\mathrm{M}$ & 29.89 \,($-0.36{\pm}0.23$) & $0.862{\pm}0.015$ & $0.0328$ & $0.269$ & $0.0118$ \\
    \bottomrule
  \end{tabular}
\end{table*}

\begin{table*}[t]
  \centering
  \small
  \caption{\textbf{Per-image sweep} (\texttt{sweep\_periimage.yml}): test-time,
           data-free methods, model-agnostic ($2$ modes $\times$ $3$ seeds;
           single row each, to be reported as mean$\pm$s.d.).
           \emph{Scaffold}---data cells (``---'') are populated as the sweep
           completes.  PSNR shows the gain over the LDCT input floor.
           $\uparrow$/$\downarrow$: higher/lower is better.}
  \label{tab:sweep-periimage}
  \begin{tabular}{llrccccc}
    \toprule
    \textbf{Method} & \textbf{Type} & \textbf{Params} &
    PSNR\,$\uparrow$ (\,$\Delta$) & SSIM\,$\uparrow$ & RMSE\,$\downarrow$ &
    GMSD\,$\downarrow$ & RSR\,$\downarrow$ \\
    \midrule
    Filter2Noise~\cite{sun2025filter2noise} & Zero-shot & --- & --- & --- & --- & --- & --- \\
    ZS-N2N~\cite{mansour2023zsn2n}          & Zero-shot & $21.3\mathrm{K}$ & --- & --- & --- & --- & --- \\
    \bottomrule
  \end{tabular}
\end{table*}

% -----------------------------------------------------------------------
\section{Discussion}

\paragraph{Architecture vs.\ objective.}
Under supervision the CNN backbones finish within $0.1$\,dB of one another
despite a $3\times$ parameter range, so the training objective dominates the
backbone at this budget.  The main benchmark sweep now completes both the
architecture and the objective comparison across three seeds
(Table~\ref{tab:sweep-benchmark}).  Two backbones underperform the CNNs even
with paired targets: CTformer gains only $+1.77$\,dB and CFM $+2.84$\,dB under
supervision, versus $\sim{+}4$\,dB for RED-CNN/U-Net/DnCNN.  Because our CFM
shares the U-Net's velocity-net body, the U-Net (MSE) vs.\ CFM (flow-matching)
contrast now isolates the objective at matched capacity: the flow-matching
objective is a $\sim$1\,dB-worse route to the supervised optimum than direct MSE
regression---a supervised-regime echo of the label-free flow-vs-regression gap
that Section~\ref{sec:method} probes directly
(Table~\ref{tab:flow-vs-reg}).

\paragraph{Metric selection.}
PSNR and SSIM penalise any pixel-level deviation, even perceptually
plausible detail, systematically disadvantaging generative models; recent
reassessments show both metrics misrank restoration algorithms on medical
images~\cite{breger2025reassess} and fail to track radiologist
perception~\cite{lee2025ldctiqa}.  GMSD correlates better with perceptual
judgements~\cite{xue2014gradient}, and RSR is directly interpretable by
radiologists as a frequency-domain characterisation of residual noise.  The
stronger standard---task-based detectability via a model
observer~\cite{li2021detection,kc2024lgcho}, together with explicit
hallucination quantification~\cite{bhadra2021halluc,tivnan2024halluc}---is
reported in Section~\ref{sec:detect}: on these benchmark models the CHO
confirms that the PSNR/SSIM gains overstate task performance, with
low-contrast detectability eroded rather than improved.  We recommend future
CT-denoising work report task-based $d'$ alongside the five metrics in
Table~\ref{tab:results}.

\paragraph{Limitations.}
The main results (Table~\ref{tab:results}) are single-split on the abdomen
window at 50 epochs; a three-fold patient-level cross-validation
(Table~\ref{tab:cv}) confirms the supervised/self-supervised ordering is stable
across patient partitions, albeit at a shorter 15-epoch budget.  The main
benchmark sweep (Table~\ref{tab:sweep-benchmark}) is complete across all five
architectures and three seeds; multi-anatomy runs and completion of the
cross-validation at the full epoch budget remain in progress.  For the proposed
self-supervised
flow (Section~\ref{sec:method}) the four-step inference arm is still completing
(ep.\,38--46), though its one- vs.\ multi-step trend is already unambiguous;
multi-seed SSFlow runs and a head-to-head at matched epochs remain to be
finalised.  The detectability evaluation (Section~\ref{sec:detect}) covers the
15-epoch cross-validation models and the deterministic denoisers; the
generative SSFlow and CFM detectability---where fabrication, not just erosion,
can arise---is in progress.

% -----------------------------------------------------------------------
\section{Conclusion}

We presented CTDenoiser, a benchmark that trains and evaluates five LDCT
denoising architectures under a shared, reproducible harness with a
corrected full-slice evaluation and two CT-appropriate metrics (GMSD,
RSR).  Once the evaluation is corrected, supervised denoisers gain
$\sim$4\,dB over the noisy input, and among label-free methods only the
correlated-noise-aware Noise2Sim reliably beats it while blind-spot
Noise2Void degrades---quantifying a $3$--$4$\,dB gap.  Our central
methodological observation is that an \emph{unconditional} rectified flow,
trained on two noisy observations of the same signal, recovers the
Noise2Noise posterior mean in a single step without clean targets; we
instantiate it as SSFlow with similarity-based pairs that decorrelate CT
noise, and it attains $+1.39$\,dB, the best label-free result in the
benchmark.  Consistent with that observation, an ablation shows the one-step
posterior mean is the operative estimator while multi-step refinement
degrades fidelity.  All code is released to facilitate
reproducibility.

% -----------------------------------------------------------------------
\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{brenner2007computed}
D.~J. Brenner and E.~J. Hall.
\newblock Computed tomography---an increasing source of radiation exposure.
\newblock \emph{New England Journal of Medicine}, 357(22):2277--2284, 2007.

\bibitem{chen2017lowdose}
H.~Chen, Y.~Zhang, M.~K. Kalra, F.~Lin, Y.~Chen, P.~Liao, J.~Zhou, and
  G.~Wang.
\newblock Low-dose {CT} with a residual encoder-decoder convolutional neural
  network.
\newblock \emph{IEEE Transactions on Medical Imaging}, 36(12):2524--2535, 2017.

\bibitem{zhang2017beyond}
K.~Zhang, W.~Zuo, Y.~Chen, D.~Meng, and L.~Zhang.
\newblock Beyond a {G}aussian denoiser: Residual learning of deep {CNN} for
  image denoising.
\newblock \emph{IEEE Transactions on Image Processing}, 26(7):3142--3155, 2017.

\bibitem{wang2022ctformer}
C.~Wang, S.~Shang, H.~Zhang, Q.~Li, and S.~K. Zhou.
\newblock {CTformer}: Convolution-free token2token dilated faster self-attention
  for {CT} image denoising.
\newblock \emph{Physics in Medicine \& Biology}, 68(6):065012, 2023.

\bibitem{moen2021low}
T.~R. Moen, B.~Chen, D.~R. Holmes, X.~Duan, Z.~Yu, L.~Yu, S.~Leng,
  J.~G. Fletcher, and C.~H. McCollough.
\newblock Low-dose {CT} image and projection data.
\newblock \emph{Medical Physics}, 48(2):902--911, 2021.

\bibitem{song2021scorebased}
Y.~Song, J.~Sohl-Dickstein, D.~P. Kingma, A.~Kumar, S.~Ermon, and B.~Poole.
\newblock Score-based generative modeling through stochastic differential
  equations.
\newblock In \emph{International Conference on Learning Representations}, 2021.

\bibitem{gao2023corediff}
Q.~Gao, F.~Li, Y.~Zhang, Y.~Wang, X.~Hu, P.~Liu, W.~Zheng, H.~Chen, and
  S.~Li.
\newblock {CoreDiff}: Contextual error-modulated generalized diffusion model for
  low-dose {CT} denoising and generalization.
\newblock \emph{IEEE Transactions on Medical Imaging}, 43(2):745--759, 2024.

\bibitem{lipman2022flow}
Y.~Lipman, R.~T.~Q. Chen, H.~Ben-Hamu, M.~Nickel, and M.~Le.
\newblock Flow matching for generative modeling.
\newblock In \emph{International Conference on Learning Representations}, 2023.

\bibitem{albergo2022building}
M.~S. Albergo and E.~Vanden-Eijnden.
\newblock Building normalizing flows with stochastic interpolants.
\newblock In \emph{International Conference on Learning Representations}, 2023.

\bibitem{liu2022flow}
X.~Liu, C.~Gong, and Q.~Liu.
\newblock Flow straight and fast: Learning to generate and transfer data with
  rectified flow.
\newblock In \emph{International Conference on Learning Representations}, 2023.

\bibitem{lehtinen2018noise2noise}
J.~Lehtinen, J.~Munkberg, J.~Hasselgren, S.~Laine, T.~Karras, M.~Aittala,
  and T.~Aila.
\newblock {Noise2Noise}: Learning image restoration without clean data.
\newblock In \emph{International Conference on Machine Learning}, 2018.

\bibitem{krull2019noise2void}
A.~Krull, T.-O. Buchholz, and F.~Jug.
\newblock {Noise2Void}---learning denoising from single noisy images.
\newblock In \emph{IEEE Conference on Computer Vision and Pattern Recognition},
  2019.

\bibitem{huang2021neighbor2neighbor}
T.~Huang, S.~Li, X.~Jia, H.~Lu, and J.~Liu.
\newblock {Neighbor2Neighbor}: Self-supervised denoising from single noisy
  images.
\newblock In \emph{IEEE Conference on Computer Vision and Pattern Recognition},
  2021.

\bibitem{niu2020noise2sim}
C.~Niu and G.~Wang.
\newblock {Noise2Sim}---similarity-based self-learning for image denoising.
\newblock \emph{arXiv preprint arXiv:2011.03384}, 2020.
\newblock Published as ``Noise suppression with similarity-based
  self-supervised deep learning,'' \emph{IEEE Transactions on Medical Imaging},
  42(6):1590--1602, 2023.

\bibitem{mansour2023zsn2n}
Y.~Mansour and R.~Heckel.
\newblock Zero-shot {Noise2Noise}: Efficient image denoising without any data.
\newblock In \emph{IEEE Conference on Computer Vision and Pattern Recognition},
  2023.

\bibitem{sun2025filter2noise}
Y.~Sun, L.-S. Schneider, M.~Gu, S.~Mei, C.~Ye, F.~Wagner, S.~Bayer, and
  A.~Maier.
\newblock {Filter2Noise}: Interpretable self-supervised single-image denoising
  for low-dose {CT} with attention-guided bilateral filtering.
\newblock \emph{arXiv preprint arXiv:2504.13519}, 2025.

% --- ADDED 2026-06: lean must-cites supporting the task-based / hallucination
% --- framing and the correlated-noise self-supervised cluster. Verify
% --- volume/page/DOI metadata flagged with TODO before submission.

\bibitem{lee2022apbsn}
W.~Lee, S.~Son, and K.~M. Lee.
\newblock {AP-BSN}: Self-supervised denoising for real-world images via
  asymmetric {PD} and blind-spot network.
\newblock In \emph{IEEE/CVF Conference on Computer Vision and Pattern
  Recognition}, pages 17725--17734, 2022.

\bibitem{zhao2024wia}
H.~Zhao, Y.~Gu, Q.~Zhao, B.~Du, X.~Xu, and Y.~Yu.
\newblock {WIA-LD2ND}: Wavelet-based image alignment for self-supervised
  low-dose {CT} denoising.
\newblock In \emph{Medical Image Computing and Computer-Assisted Intervention
  (MICCAI)}, LNCS vol.\ 15007, 2024.
% TODO verify page range.

\bibitem{hein2025pfcm}
D.~Hein, G.~Stevens, A.~Wang, and G.~Wang.
\newblock {PFCM}: Poisson flow consistency models for low-dose {CT} image
  denoising.
\newblock \emph{IEEE Transactions on Medical Imaging}, 44(7):2989--3001, 2025.

\bibitem{blau2018perception}
Y.~Blau and T.~Michaeli.
\newblock The perception-distortion tradeoff.
\newblock In \emph{IEEE Conference on Computer Vision and Pattern Recognition},
  pages 6228--6237, 2018.

\bibitem{bhadra2021halluc}
S.~Bhadra, V.~A. Kelkar, F.~J. Brooks, and M.~A. Anastasio.
\newblock On hallucinations in tomographic image reconstruction.
\newblock \emph{IEEE Transactions on Medical Imaging}, 40(11):3249--3260, 2021.

\bibitem{tivnan2024halluc}
M.~Tivnan, S.~Yoon, Z.~Chen, X.~Li, D.~Wu, and Q.~Li.
\newblock Hallucination index: An image quality metric for generative
  reconstruction models.
\newblock In \emph{Medical Image Computing and Computer-Assisted Intervention
  (MICCAI)}, LNCS vol.\ 15010, pages 449--458, 2024.

\bibitem{li2021detection}
K.~Li, W.~Zhou, H.~Li, and M.~A. Anastasio.
\newblock Assessing the impact of deep neural network-based image denoising on
  binary signal detection tasks.
\newblock \emph{IEEE Transactions on Medical Imaging}, 40(9):2295--2305, 2021.

\bibitem{kc2024lgcho}
P.~Kc and R.~Zeng.
\newblock Assessing the performance of {CT} image denoisers using
  {Laguerre-Gauss} channelized {Hotelling} observer for lesion detection.
\newblock In \emph{IEEE Nuclear Science Symposium and Medical Imaging Conference
  (NSS/MIC)}, 2024.
\newblock arXiv:2412.02920.

\bibitem{eulig2024benchmark}
E.~Eulig, B.~Ommer, and M.~Kachelrie\ss.
\newblock Benchmarking deep learning-based low-dose {CT} image denoising
  algorithms.
\newblock \emph{Medical Physics}, 51(12):8776--8788, 2024.

\bibitem{breger2025reassess}
A.~Breger, A.~Biguri, M.~S. Saba\'t\'e Landman, I.~Selby, N.~Amberg,
  E.~Brunner, J.~Gr\"ohl, S.~Hatamikia, C.~Karner, L.~Ning, S.~Dittmer,
  M.~Roberts, AIX-COVNET Collaboration, and C.-B. Sch\"onlieb.
\newblock A study of why we need to reassess full reference image quality
  assessment with medical images.
\newblock \emph{Journal of Imaging Informatics in Medicine},
  38(6):3444--3469, 2025.

\bibitem{lee2025ldctiqa}
W.~Lee, F.~Wagner, A.~Galdran, et~al.
\newblock Low-dose computed tomography perceptual image quality assessment.
\newblock \emph{Medical Image Analysis}, 99:103343, 2025.

\bibitem{wang2004image}
Z.~Wang, A.~C. Bovik, H.~R. Sheikh, and E.~P. Simoncelli.
\newblock Image quality assessment: From error visibility to structural
  similarity.
\newblock \emph{IEEE Transactions on Image Processing}, 13(4):600--612, 2004.

\bibitem{xue2014gradient}
W.~Xue, L.~Zhang, X.~Mou, and A.~C. Bovik.
\newblock Gradient magnitude similarity deviation: A highly efficient perceptual
  image quality index.
\newblock \emph{IEEE Transactions on Image Processing}, 23(2):684--695, 2014.

\bibitem{zhang2018unreasonable}
R.~Zhang, P.~Isola, A.~A. Efros, E.~Shechtman, and O.~Wang.
\newblock The unreasonable effectiveness of deep features as a perceptual
  metric.
\newblock In \emph{IEEE Conference on Computer Vision and Pattern Recognition},
  2018.

\bibitem{solomon2012characterization}
J.~B. Solomon, O.~Christianson, and E.~Samei.
\newblock Quantitative comparison of noise texture across {CT} scanners from
  different manufacturers.
\newblock \emph{Medical Physics}, 39(10):6048--6055, 2012.

\end{thebibliography}

\end{document}