\documentclass[10pt,twocolumn]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{microtype}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!70!black,
  citecolor=blue!70!black,
  urlcolor=blue!70!black,
}

% -----------------------------------------------------------------------
\title{\textbf{CTDenoiser}: A Benchmark for Low-Dose CT Denoising\\
       Across Supervised and Self-Supervised Regimes}

\author{Tim Sereda}
% -----------------------------------------------------------------------

\begin{document}
\maketitle

% -----------------------------------------------------------------------
\begin{abstract}
Low-dose computed tomography (LDCT) denoising must contend with
structured, spatially correlated noise that ordinary blind-spot
self-supervision is not designed to handle.  We are building
\textbf{CTDenoiser}, a benchmark and single data/training/inference harness
for comparing denoisers across supervision regimes---fully
\emph{supervised} training with paired full-dose targets,
\emph{Noise2Void} blind-spot self-supervision, \emph{Noise2Sim}
similarity-based self-supervision, and \emph{Zero-Shot Noise2Noise}
(ZS-N2N) per-image test-time optimisation---reported with PSNR, SSIM, and
RMSE alongside Gradient Magnitude Similarity Deviation (GMSD) and a Noise
Power Spectrum ratio (NPS-ratio) that targets spurious high-frequency
texture.  This is an interim report.  We present two complementary sweeps: an
\emph{architecture} sweep comparing five backbones under supervised and
Noise2Void training alongside zero-shot ZS-N2N, and a
\emph{supervision-regime} sweep on the RED-CNN and DnCNN backbones spanning
supervised, Noise2Sim, and Noise2Void training; several self-supervised
runs are still in progress.  We recently identified and fixed an
evaluation bug in overlapped full-slice inference that zeroed each slice's
outer border and so biased reconstruction metrics against \emph{every}
trained model; the numbers below predate that fix and are being
regenerated.  Even under the biased harness, the structural-similarity
ordering across methods is sensible and supervised RED-CNN already improves
SSIM over the noisy input.
\end{abstract}

% -----------------------------------------------------------------------
\section{Introduction}
Reducing X-ray exposure lowers patient radiation dose~\cite{brenner2007computed}
but raises image noise that is spatially correlated by the reconstruction
process, unlike the pixel-independent noise assumed by many learned
denoisers.  Our goal is a reproducible benchmark that places supervised,
self-supervised, and zero-shot denoisers on common footing---identical
data splits, preprocessing, training budget, full-slice inference, and
metrics---so that differences reflect the methods rather than the harness.
We build on the public TCIA LDCT-and-Projection-Data
collection~\cite{moen2021low}.

% -----------------------------------------------------------------------
\section{Related Work}

\paragraph{CNN-based denoisers.}
RED-CNN~\cite{chen2017lowdose} introduced a residual encoder-decoder with
multi-level skip connections and remains a standard baseline.
DnCNN~\cite{zhang2017beyond} showed that a flat stack of Conv-BN-ReLU
layers predicting the residual (noise) transfers well across denoising
tasks.

\paragraph{Transformer and generative denoisers.}
CTformer~\cite{wang2022ctformer} applies a U-shaped transformer with
Token2Token dilation to capture long-range structure at reduced memory
cost.  Score-based diffusion~\cite{song2021scorebased,gao2023corediff}
can yield perceptually strong results but needs many neural function
evaluations; conditional flow matching~\cite{lipman2022flow,albergo2022building}
learns near-straight transport paths and cuts this to tens of steps.  These
sit on our roadmap but are not yet evaluated here.

\paragraph{Self-supervised and zero-shot denoisers.}
When paired full-dose targets are unavailable, denoisers can train from
noisy data alone.  Noise2Noise~\cite{lehtinen2018noise2noise} maps between
two noisy observations of a scene; Noise2Void~\cite{krull2019noise2void}
removes the second observation via a blind-spot scheme but assumes
pixel-wise \emph{independent} noise---an assumption CT violates.  Two lines
of work target correlated noise directly:
Noise2Sim~\cite{niu2020noise2sim} maps between centre pixels of
self-similar non-local patches, and Zero-Shot Noise2Noise
(ZS-N2N)~\cite{mansour2023zsn2n} trains a tiny per-image network from a
downsampled Noise2Noise pair with no external data.  More recently,
Filter2Noise~\cite{sun2025filter2noise} couples single-image
self-supervision with an interpretable attention-guided bilateral filter
for LDCT.  Our harness includes Noise2Void and ZS-N2N as blind-spot and
zero-shot baselines and adds Noise2Sim as the correlated-noise-aware
extension; Filter2Noise is integrated and currently running.

\paragraph{Evaluation metrics.}
PSNR and SSIM~\cite{wang2004image} favour pixel-level fidelity and can
penalise plausible-but-incorrect high-frequency detail.  We complement
them with GMSD~\cite{xue2014gradient} as an efficient perceptual proxy and
an NPS-ratio inspired by noise-texture
characterisation~\cite{solomon2012characterization} to flag spurious
residual texture; LPIPS~\cite{zhang2018unreasonable} is a planned addition.

% -----------------------------------------------------------------------
\section{Methods}

\paragraph{Data and preprocessing.}
We use a subset of the TCIA LDCT-and-Projection-Data~\cite{moen2021low}
with a patient-level split (seed 0): 10 training patients
($6{,}683$ slices) and a held-out validation set ($1{,}476$ slices).
HU values are windowed to $[0,1]$ by $\mathrm{clip}((\mathrm{HU}+b)/s)$
with a fixed anatomy preset.

\paragraph{Architectures and supervision.}
The harness supports RED-CNN, DnCNN, U-Net, CTformer, and a conditional
flow-matching model (CFM), trained under five modes---supervised,
Noise2Void, Noise2Sim, ZS-N2N, and Filter2Noise.  Table~\ref{tab:arch}
sweeps architectures under supervised/N2V training; Table~\ref{tab:results}
sweeps supervision regimes on the RED-CNN and DnCNN backbones.

\paragraph{Inference and evaluation.}
Trained models are applied to full slices by overlapped tiling.  We score
every model against an identity ``do-nothing'' baseline (the noisy input)
so each row is read as a gain over that floor.

\paragraph{Evaluation fix (in progress).}
We found that the overlapped-tiling routine discarded a fixed margin from
\emph{every} patch, including patches at the image edge, leaving the outer
border of each output slice at zero.  Because the identity baseline does
not pass through this routine, the artefact penalised only the trained
models and biased their PSNR/RMSE pessimistically.  The fix retains the
margin on sides that touch the image edge; runs are being regenerated.
\textbf{Table~\ref{tab:results} predates this fix.}

% -----------------------------------------------------------------------
\section{Preliminary Results}

We report two sweeps under the shared harness.  Both predate the
overlapped-inference border fix (Sec.~3) and so are biased pessimistically
on the reconstruction metrics (PSNR/RMSE); a corrected re-run is in
progress.  We use them here for \emph{relative} comparison, where the
ordering is robust to that bias.

\paragraph{Architecture sweep.}
Table~\ref{tab:arch} fixes the supervision mode and varies the backbone.
Under supervised training the convolutional baselines (RED-CNN, DnCNN,
U-Net) cluster tightly and lead on SSIM, while the zero-shot ZS-N2N---which
trains a $21$K-parameter network per image with no external data---is
strongest on the noise-texture metrics (NPS-ratio, GMSD), as expected for a
method that barely departs from the input.

\paragraph{Supervision-regime sweep.}
Table~\ref{tab:results} fixes the backbone (RED-CNN, DnCNN) and varies the
training regime.  The ordering is the expected one---supervised $>$
Noise2Sim $>$ Noise2Void on SSIM---with the correlated-noise-aware
Noise2Sim ahead of the blind-spot Noise2Void, consistent with CT noise
violating the pixel-independence assumption.  Supervised RED-CNN improves
SSIM over the noisy input ($0.924$ vs.\ $0.893$), evidence it recovers
structure rather than merely smoothing.

\begin{table*}[t]
  \centering
  \caption{\textbf{Architecture sweep.}  Validation results on the TCIA
           LDCT-and-Projection-Data benchmark (patient-level $80/20$ split,
           seed 0; trained models at 50 epochs).  Means over the validation
           set on full slices via overlapped tiling.  $\uparrow$/$\downarrow$:
           higher/lower is better.  \emph{LDCT input} is the noisy scan with
           no denoising.  \textbf{Bold}: best denoiser per metric.  ZS-N2N
           values are mean\,$\pm$\,std over its five model-agnostic runs; all
           other rows are single runs (seed 0).  Numbers predate the
           border-fix of Sec.~3.}
  \label{tab:arch}
  \begin{tabular}{llrccccc}
    \toprule
    \textbf{Model} & \textbf{Supervision} & \textbf{Params} &
    PSNR\,$\uparrow$ & SSIM\,$\uparrow$ & RMSE\,$\downarrow$ &
    GMSD\,$\downarrow$ & NPS-r.\,$\downarrow$ \\
    \midrule
    LDCT input & --- (no denoising) & --- & 30.03 & 0.883 & 0.0331 & 0.171 & 0.0122 \\
    \midrule
    RED-CNN~\cite{chen2017lowdose}   & Supervised & $1.85\mathrm{M}$ & 27.93 & \textbf{0.913} & 0.0413 & 0.204 & 0.0195 \\
    DnCNN~\cite{zhang2017beyond}     & Supervised & $0.56\mathrm{M}$ & 27.88 & 0.908 & 0.0416 & 0.209 & 0.0197 \\
    U-Net                            & Supervised & $1.95\mathrm{M}$ & 27.82 & 0.905 & 0.0418 & 0.226 & 0.0198 \\
    CTformer~\cite{wang2022ctformer} & Supervised & $0.40\mathrm{M}$ & 27.48 & 0.850 & 0.0434 & 0.296 & 0.0212 \\
    CFM (ours)                       & Supervised & $2.11\mathrm{M}$ & 27.38 & 0.887 & 0.0440 & 0.245 & 0.0219 \\
    \midrule
    RED-CNN~\cite{chen2017lowdose}   & Noise2Void & $1.85\mathrm{M}$ & 26.63 & 0.875 & 0.0478 & \textbf{0.201} & 0.0255 \\
    DnCNN~\cite{zhang2017beyond}     & Noise2Void & $0.56\mathrm{M}$ & 26.51 & 0.870 & 0.0485 & 0.232 & 0.0262 \\
    U-Net                            & Noise2Void & $1.95\mathrm{M}$ & 26.67 & 0.875 & 0.0475 & 0.211 & 0.0251 \\
    CTformer~\cite{wang2022ctformer} & Noise2Void & $0.40\mathrm{M}$ & 26.15 & 0.857 & 0.0503 & 0.298 & 0.0280 \\
    CFM                              & Noise2Void & --- & \multicolumn{5}{c}{\textit{undefined (requires paired targets)}} \\
    \midrule
    ZS-N2N~\cite{mansour2023zsn2n}   & Zero-shot & $21.3\mathrm{K}$ &
      \textbf{29.92}{\scriptsize\,$\pm0.001$} & 0.882{\scriptsize\,$\pm0.0005$} &
      \textbf{0.0331} & \textbf{0.188}{\scriptsize\,$\pm0.0001$} & \textbf{0.0121} \\
    \bottomrule
  \end{tabular}
\end{table*}

\begin{table*}[t]
  \centering
  \caption{\textbf{Supervision-regime sweep.}  Validation results on a TCIA
           LDCT-and-Projection-Data subset (patient-level split, seed 0;
           $10$ training epochs).  Means over the validation set on full
           slices via overlapped tiling.  \textbf{Bold}: best denoiser per
           metric.  Rows marked $^{\dagger}$ are runs still in progress
           (epoch shown); ZS-N2N and Filter2Noise runs had not logged
           validation metrics at export time.  Numbers predate the
           border-fix of Sec.~3.}
  \label{tab:results}
  \begin{tabular}{llrccccc}
    \toprule
    \textbf{Model} & \textbf{Supervision} & \textbf{Params} &
    PSNR\,$\uparrow$ & SSIM\,$\uparrow$ & RMSE\,$\downarrow$ &
    GMSD\,$\downarrow$ & NPS-r.\,$\downarrow$ \\
    \midrule
    LDCT input & --- (no denoising) & --- & 31.44 & 0.893 & 0.0274 & 0.166 & 0.0079 \\
    \midrule
    RED-CNN~\cite{chen2017lowdose} & Supervised & $1.85\mathrm{M}$ & \textbf{27.99} & \textbf{0.924} & \textbf{0.0408} & 0.200 & \textbf{0.0179} \\
    RED-CNN~\cite{chen2017lowdose} & Noise2Sim~\cite{niu2020noise2sim} & $1.85\mathrm{M}$ & 27.28 & 0.896 & 0.0441 & \textbf{0.179} & 0.0208 \\
    RED-CNN~\cite{chen2017lowdose} & Noise2Void~\cite{krull2019noise2void} & $1.85\mathrm{M}$ & 26.95 & 0.882 & 0.0457 & 0.195 & 0.0222 \\
    \midrule
    DnCNN~\cite{zhang2017beyond} & Supervised$^{\dagger}$\,(ep.\,8) & $0.56\mathrm{M}$ & 27.96 & 0.922 & 0.0409 & 0.206 & 0.0180 \\
    DnCNN~\cite{zhang2017beyond} & Noise2Sim$^{\dagger}$\,(ep.\,6) & $0.56\mathrm{M}$ & 27.35 & 0.843 & 0.0438 & 0.223 & 0.0205 \\
    DnCNN~\cite{zhang2017beyond} & Noise2Void$^{\dagger}$\,(ep.\,6) & $0.56\mathrm{M}$ & 26.82 & 0.878 & 0.0463 & 0.235 & 0.0228 \\
    \bottomrule
  \end{tabular}
\end{table*}

% -----------------------------------------------------------------------
\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{brenner2007computed}
D.~J. Brenner and E.~J. Hall.
\newblock Computed tomography---an increasing source of radiation exposure.
\newblock \emph{New England Journal of Medicine}, 357(22):2277--2284, 2007.

\bibitem{chen2017lowdose}
H.~Chen, Y.~Zhang, M.~K. Kalra, F.~Lin, Y.~Chen, P.~Liao, J.~Zhou, and
  G.~Wang.
\newblock Low-dose {CT} with a residual encoder-decoder convolutional neural
  network.
\newblock \emph{IEEE Transactions on Medical Imaging}, 36(12):2524--2535, 2017.

\bibitem{zhang2017beyond}
K.~Zhang, W.~Zuo, Y.~Chen, D.~Meng, and L.~Zhang.
\newblock Beyond a {G}aussian denoiser: Residual learning of deep {CNN} for
  image denoising.
\newblock \emph{IEEE Transactions on Image Processing}, 26(7):3142--3155, 2017.

\bibitem{wang2022ctformer}
C.~Wang, S.~Shang, H.~Zhang, Q.~Li, and S.~K. Zhou.
\newblock {CTformer}: Convolution-free token2token dilated faster self-attention
  for {CT} image denoising.
\newblock \emph{Physics in Medicine \& Biology}, 68(6):065012, 2023.

\bibitem{moen2021low}
T.~R. Moen, B.~Chen, D.~R. Holmes, X.~Duan, Z.~Yu, L.~Yu, S.~Leng,
  J.~G. Fletcher, and C.~H. McCollough.
\newblock Low-dose {CT} image and projection data.
\newblock \emph{Medical Physics}, 48(2):902--911, 2021.

\bibitem{song2021scorebased}
Y.~Song, J.~Sohl-Dickstein, D.~P. Kingma, A.~Kumar, S.~Ermon, and B.~Poole.
\newblock Score-based generative modeling through stochastic differential
  equations.
\newblock In \emph{International Conference on Learning Representations}, 2021.

\bibitem{gao2023corediff}
Q.~Gao, F.~Li, Y.~Zhang, Y.~Wang, X.~Hu, P.~Liu, W.~Zheng, H.~Chen, and
  S.~Li.
\newblock {CoreDiff}: Contextual error-modulated generalized diffusion model for
  low-dose {CT} denoising and generalization.
\newblock \emph{IEEE Transactions on Medical Imaging}, 43(2):745--759, 2024.

\bibitem{lipman2022flow}
Y.~Lipman, R.~T.~Q. Chen, H.~Ben-Hamu, M.~Nickel, and M.~Le.
\newblock Flow matching for generative modeling.
\newblock In \emph{International Conference on Learning Representations}, 2023.

\bibitem{albergo2022building}
M.~S. Albergo and E.~Vanden-Eijnden.
\newblock Building normalizing flows with stochastic interpolants.
\newblock In \emph{International Conference on Learning Representations}, 2023.

\bibitem{lehtinen2018noise2noise}
J.~Lehtinen, J.~Munkberg, J.~Hasselgren, S.~Laine, T.~Karras, M.~Aittala,
  and T.~Aila.
\newblock {Noise2Noise}: Learning image restoration without clean data.
\newblock In \emph{International Conference on Machine Learning}, 2018.

\bibitem{krull2019noise2void}
A.~Krull, T.-O. Buchholz, and F.~Jug.
\newblock {Noise2Void}---learning denoising from single noisy images.
\newblock In \emph{IEEE Conference on Computer Vision and Pattern Recognition},
  2019.

\bibitem{niu2020noise2sim}
C.~Niu and G.~Wang.
\newblock {Noise2Sim}---similarity-based self-learning for image denoising.
\newblock \emph{arXiv preprint arXiv:2011.03384}, 2020.

\bibitem{mansour2023zsn2n}
Y.~Mansour and R.~Heckel.
\newblock Zero-shot {Noise2Noise}: Efficient image denoising without any data.
\newblock In \emph{IEEE Conference on Computer Vision and Pattern Recognition},
  2023.

\bibitem{sun2025filter2noise}
Y.~Sun, L.-S. Schneider, M.~Gu, S.~Mei, C.~Ye, F.~Wagner, S.~Bayer, and
  A.~Maier.
\newblock {Filter2Noise}: Interpretable self-supervised single-image denoising
  for low-dose {CT} with attention-guided bilateral filtering.
\newblock \emph{arXiv preprint arXiv:2504.13519}, 2025.

\bibitem{wang2004image}
Z.~Wang, A.~C. Bovik, H.~R. Sheikh, and E.~P. Simoncelli.
\newblock Image quality assessment: From error visibility to structural
  similarity.
\newblock \emph{IEEE Transactions on Image Processing}, 13(4):600--612, 2004.

\bibitem{xue2014gradient}
W.~Xue, L.~Zhang, X.~Mou, and A.~C. Bovik.
\newblock Gradient magnitude similarity deviation: A highly efficient perceptual
  image quality index.
\newblock \emph{IEEE Transactions on Image Processing}, 23(2):684--695, 2014.

\bibitem{zhang2018unreasonable}
R.~Zhang, P.~Isola, A.~A. Efros, E.~Shechtman, and O.~Wang.
\newblock The unreasonable effectiveness of deep features as a perceptual
  metric.
\newblock In \emph{IEEE Conference on Computer Vision and Pattern Recognition},
  2018.

\bibitem{solomon2012characterization}
J.~B. Solomon, O.~Christianson, and E.~Samei.
\newblock Quantitative comparison of noise texture across {CT} scanners from
  different manufacturers.
\newblock \emph{Medical Physics}, 39(10):6048--6055, 2012.

\end{thebibliography}

\end{document}
